---
title: "LFS2014"
author: "Jian Hao, Isanna"
date: ''
output:
  pdf_document: default
  html_document: default
subtitle: OLS
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Set global R options
#options("scipen"=10, "digits"=4)

# kable
options(knitr.kable.NA = '')

# Set the graphical theme
ggplot2::theme_set(ggplot2::theme_light())

# load libraries -- install if required
library(pacman)
p_load(tidyverse,
       dplyr,
       oaxaca,
       knitr,
       ggplot2,
       haven,
       Hmisc,
       survival,
       Formula,
       lattice,
       matrixStats,
       fastDummies,
       magrittr)
      
```

<div class="alert alert-info">
  <strong>Group Members:</strong> 
1. Jian Hao Chiah, 216031965
2. Isanna Biglands, 218887273
</div>

# Part 1: Importing Data & Data Manipulation
```{r, echo=TRUE}
A14 <- read.csv("D:/Desktop/PROC A/Data/2014/1.csv") 
B14 <- read.csv("D:/Desktop/PROC A/Data/2014/2.csv") 

raw <- rbind(A14, B14)
```

```{r eval=T, echo=T}
#Cleaning the data -- Part 1

lfs <- raw %>% 
  drop_na(FTPTMAIN) %>% 
  drop_na(HRLYEARN) %>%
  drop_na(PERMTEMP) %>%
  mutate(hrlywage = HRLYEARN/100,
         hours = UTOTHRS/10,
         average_weekly_wage = hrlywage*hours,
         lWAGE = log(average_weekly_wage),
         FEMALE = ifelse(SEX == 2, 1, 0),
         full_part = ifelse(FTPTMAIN == 2, 1, 0),
         perm_temp = ifelse(PERMTEMP == 2, 1, 0)) 
```
```{r eval=T, echo=T}
#Creating Dummies with more than 2 categories
lfs <- dummy_cols(lfs, select_columns = c('AGE_12','MARSTAT','EDUC90','CMA'),
                  ignore_na = FALSE)

```
```{r eval=T, echo=T}
df <- lfs %>% 
  select(AGE_12, FEMALE, lWAGE, MARSTAT, EDUC90, MJH, NAICS_18, UHRSMAIN, UTOTHRS, TENURE, UNION, SCHOOLN, AGYOWNKN, FWEIGHT)

FEMALE.f <- factor(df$FEMALE, labels=c("men", "women"))

```

# Part 2: Overlapping Density Plots based on Gender
```{r eval=T, echo=T, results='hide'}
# Log Wage by Gender
df %>% 
  mutate(FEMALE.f <- factor(FEMALE, labels = c("MEN","WOMEN")))
```
## Log Wage by Gender
```{r eval=T, echo=T}
ggplot(df, aes(x = lWAGE, y = ..density.., weight = FWEIGHT, fill = FEMALE.f)) +
         geom_histogram(bins = 30, position = position_dodge(width = 0), alpha = 1) +
  labs(x = "Log Wage", y = "Density", title = "Log(wage) by Gender") +
  scale_x_log10() +
  theme_classic()
```

## Highest Educational Attainment by Gender
```{r eval=T, echo=T}
# Highest Educational Attainment by Gender
ggplot(df, aes(x = EDUC90, y = ..density.., weight = FWEIGHT, fill = FEMALE.f)) +
  geom_histogram(bins = 20, position = position_dodge(width = 0), alpha = 1) +
  labs(x = "Highest Education Attainment", y = "Density", title = "Highest Educational Attainment by Gender") +
  theme_classic()

```

# Part 3: Average wage differences
## Avg.wage diff. by Gender
```{r eval=T, echo=T}
df.gen <- df %>% 
  group_by(FEMALE) %>% 
  summarise(mwage = weighted.mean(lWAGE, FWEIGHT),
            median.wage = weightedMedian(lWAGE, FWEIGHT))

df.gen
```
## Avg.wage diff. by Highest Educational Attainment
```{r eval=T, echo=T}
df.educ <- df %>% 
  group_by(EDUC90) %>% 
  summarise(mean.wage = weighted.mean(lWAGE, FWEIGHT),
            median.wage = weightedMedian(lWAGE, FWEIGHT))

df.educ
```
# Part 4: Standard deviation for Log Wages
## Std. Dev. for Log Wages by Gender
```{r eval=T, echo=T}
df.gen <- df %>% 
  group_by(FEMALE) %>% 
  summarise(std.dev.wage = sd(lWAGE, na.rm = F))

df.gen
```
## Std. Dev. for Log Wages by Highest Educational Attainment
```{r eval=T, echo=T}
df.gen <- df %>% 
  group_by(EDUC90) %>% 
  summarise(std.dev.wage = sd(lWAGE, na.rm = F))

df.gen
```
# Part 5: Pooled Weighted OLS model
```{r eval=T, echo=T}
OLS <- lm(lWAGE ~ FEMALE
             + AGE_12_1 
             + AGE_12_2 
             + AGE_12_3
             + AGE_12_4
             + AGE_12_5
             + AGE_12_6
             + AGE_12_7
             + AGE_12_8
             + AGE_12_9
             + AGE_12_10
             + AGE_12_11
             + AGE_12_12
             + MARSTAT_1
             + MARSTAT_2
             + MARSTAT_3
             + MARSTAT_4
             + MARSTAT_5
             + MARSTAT_6
             + EDUC90_1
             + EDUC90_2
             + EDUC90_3
             + EDUC90_4
             + EDUC90_5
             + EDUC90_6
             + CMA_1
             + CMA_1
             + CMA_2
             + CMA_3
             + CMA_4
             + full_part , weight = FWEIGHT, data = lfs)

summary(OLS)
  
```
```{r eval=T, echo=T}
summary(OLS)$r.squared
```
# Part 6: Oaxaca-Blinder Decomposition Method based on Gender
```{r eval=F, echo=T, results='hide'}
#explain differences in log real wages across gender
OB <- oaxaca(formula = lWAGE ~ AGE_12_1 
             + AGE_12_2 
             + AGE_12_3
             + AGE_12_4
             + AGE_12_5
             + AGE_12_6
             + AGE_12_7
             + AGE_12_8
             + AGE_12_9
             + AGE_12_10
             + AGE_12_11
             + AGE_12_12
             + MARSTAT_1
             + MARSTAT_2
             + MARSTAT_3
             + MARSTAT_4
             + MARSTAT_5
             + MARSTAT_6
             + EDUC90_1
             + EDUC90_2
             + EDUC90_3
             + EDUC90_4
             + EDUC90_5
             + EDUC90_6
             + CMA_1
             + CMA_1
             + CMA_2
             + CMA_3
             + CMA_4
             + full_part + perm_temp | FEMALE, data = lfs, R = 100)

print(OB)

```